### 2021-05-03 excerpt-formalism.tex
make file structure match paper structure

\documentclass[acmsmall,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}

\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2021}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

\setcopyright{none}

\input{config-tex}
\input{config-tikz}
\input{config-glyphs}
\input{config-definitions}
\input{config-graphs}

\begin{document}

\title[Grove]{Grove: A Convergent Collaborative Structure-Editor Calculus}
\subtitle{Formalism}

\maketitle

\input{appendix-B-formalism}

\end{document}