### 2022-03-07 appendix/language-definitions.tex
simplify section file names and start separate appendix doc

\section{Language Definitions}
\label{sec:language-definitions}