### 2022-08-09 appendix/language-construction.tex
snapshot of current working directory

\section{Language Construction}
\label{sec:language-construction}

\begin{definition}
  A \emph{signature}
  $\Sig = (\Sorts, \Constructors, \Positions, \Sp, \Sc, \ArityInv)$ consists of
  \begin{enumerate}
    \item a set $\Sorts$ of \emph{sorts},
    \item a set $\Constructors$ of \emph{constructors}, or vertex types,
    \item a set $\Positions$ of \emph{positions}, or edge types,
    \item a function $\Sc : \Constructors \to \Sorts$ assigning each constructor a sort,
    \item a function $\Sp : \Positions \to \Sorts$ assigning each position a sort,
    \item a function $\ArityInv : \Positions \to \Constructors$ assigning each position a constructor.
  \end{enumerate}
\end{definition}

\begin{definition}
  The \emph{ports} $\Ports{C}$ over constructor $C$ is the set
  \[ \Setof{P \in \Positions \suchthat \ArityInv(P) = C} \]
  of positions assigned to constructor $C$.
\end{definition}

\begin{definition}
  The \emph{arity map} $\Arity : \Constructors \to \Pow{\Positions}$ of signature $\Sig$ is
  \[ \Arity = (\ArityInv)^{-1} \]
  the inverse of its constructor assignment function.
\end{definition}