### 2022-08-23 introduction.tex
intro

\section{Introduction}%
\label{sec:Introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Programming is an increasingly collaborative activity.
Today, collaborative editing services are either
provided directly by an asynchronous diff-and-merge version control system like git, darcs, or pijul,
or they are integrated into a real-time editor backed by a distributed collaboration protocol such as OT or CRDT.
Whereas algorithms for collaborative text editing have been well developed in the OT setting,
we know far less about collaborative editing of tree-structured documents.

Diff-and-merge is clearly not an ideal solution for collaborative tree editing.
Since diffs do not provide information on the original scope or ordering of edits,
the merging process must reconstruct an approximation, e.g., as a sequence of line additions and deletions.
In a tree-native setting, line-based merging can fail in ways that methods which retain finer grained information would not.
Moreover, user intent can be impossible to preserve,
so critical features such as identity-preserving repositioning are impossible to implement reliably on top of them.

Operational Transformation (OT), on the other hand, goes perhaps too far in the opposite direction.
OT is a distributed system architecture that coordinates the proliferation and maintainance of edits and edit states.
Industrial-scale collaborative editors, such as Google Docs,
employ OT for real-time applications requiring strong eventual consistency of causally ordered events.
It does so by maintaining an event log it transforms upon each new entry in order to preserve the causal ordering.

The Convergent Replicated Data Type (CRDT) was proposed as a simpler alternative to the OT architecture.
CRDTs do not require causal ordering maintenance because the language of edits must form a join semilattice,
thereby ensuring the edit state update function increases monotonically.
CRDT places fewer constraints than OT on the language of edits,
so calculi for editing tree-native documents can be simpler to model for CRDT than OT.
Similarly, by eliminating all but one design constraint, edit calculi for CRDT can be easier to implement and extend.

Unfortunately, monotonicity merely changes the nature of the complexity rather than eliminating it entirely.
Although tree-native document editing calculi may ultimately admit simpler designs under CRDT than OT,
there are no generic recipes for transforming a sequential tree editing calculus into one suitable for CRDT.
Consequently, progress of CRDT-backed tree editors has slowed as the research community debates whether CRDT is living up to its promise.
Thus, we are motivated to ask:
Is there a more natural model for collaborative editing of trees?
And how do we preserve user intent in the presence of concurrent edits in a tree-native editor?

To address these gaps in the literature, we propose \emph{Grove}, a CRDT-backed calculus for collaborative structure editing.
At its heart is a simple partially ordered set of operations over a labeled multidigraph with a distinguished root vertex.
In this setting, conflicting edit states may be reified as multi-edges,
allowing conflicts to be exposed as they arise and enabling novel forms of collaboration.
We establish an isomorphism between labeled multidigraphs and a collection of forests called a \emph{grove},
enabling more familiar interactions with sub-trees of the graph in more familiar settings, e.g., as a term in a programming language editor.

We are further motivated by whether there exists a more natural medium for collaborative extension of sophisticated structure editors like Hazel.
Hazel maintains a sensibility invariant, i.e., every edit state is semantically valid.
Scaling editor services in the presence of many collaborators is difficult but necessary in a world of increasing connectivity.
By providing a first step in that direction,
we believe Grove has the potential to spark a new wave of inquiry into collaborative structure editing.

% algorithms for collaborative text editing are well developed
% natural fit for text
% not a natural fit for trees
% so if you want to build a grove, a more natural approach is needed

% a lot of collaborative editing of other sorts (not necessarily programming) requires a tree model,
% but that problem doesn't seem to have been study directly (maybe except JSON CRDT).

% A \emph{collaborative structure editor} is an editor that
% (1) allows multiple concurrent users to work on a shared document, while also
% (2) providing structure-aware editor services such as projectional editing, syntax highlighting, or automatic code folding.
% %
% Collaborative editing research focuses on the design and implementation of real-time, multi-user, character-based communication systems,
% whereas structure editors typically presume a more complex document schema and then focus on some other aspect of the user experience.
% In both settings, preservation of user intent is a core technical challenge.
% %
% Although collaborative editors and structure editors have overlapping goals (optimal user experience)
% and complementary design challenges (subject-subject versus subject-object harmony),
% to our knowledge, there is no comprehensive, principled account of their combined use.

% Since collaborative editors are essentially distributed systems, existing work tends to focus on extensions to distribution protocols.
% Lots of examples using OT. (Google Docs)
% OT is complex and largely textual.
% OT can make sense for real time systems: users typically change one character at a time, and instant feedback can help to prevent conflicts.
% On the other hand, OT system designs can be difficult to extend.

% Alternatively, there's CRDT. (Peritext?)
% CRDT is easier to implement, but harder to design for.
% There's an All-CRDT editor---it turned out to be not so realistic. (what's it called again?)

% Structure editing has been a recurring theme in the computer science literature since at least Engelbart's ``Mother of All Demos.''
% Provides automation for domain experts and reduces the barrier to entry for everyone else.
% Popular for editing programs, i.e., for programming language-specific editors.

% However, modern program editors typically disable editor services (like what?) when the document is not in a consistent state,
% a phenomenon called the ``gap problem.''
% Of course, in the presence of multiple concurrent users, the problem gets worse.

% In a collaborative setting,
% Hazel is a structure editor with support for advanced editing services, (e.g., semantic actions?).

\newpage

Motivation:

- collaborative editing (both synchronous ala Google Docs and asynchronous version control)
is good and important as computing grows

- semantic structure editing is good because it solves the gap problem (semantic editor services
are always available) -- cite Hazelnut papers (talk about holes)

- previous approaches to collaborative editing have limitations

- diff/merge based approaches (trying to solve the inverse problem based on final states --
you lose the actual actions that were performed, and have to reconstruct them or an approx.
of them i.e. add line/delete line actions -- would need to adapt this to structure editing,
some papers have started to look at that, but fundamentally we don't want to throw away the
knowledge we have about the edits!)

- operational transforms (complexity, you have to patch previous actions based on new actions)

- CRDT-based collaborative editing (that's all been on text, not PL semantics) -- this is good
because it is relatively simple: you just send all the edits to all the replicas and they are
convergent by design

- we want to have the same convergence for a CRDT-based collaborative structure editor that maintains
the sensibility invariant of Hazelnut, i.e. every editor state has meaning. mention that maintaining sensibility
allows scaling of semantic editor services in the presence of large number of collaborators (in contrast,
using VS Code or other collaborative text editors with large numbers of collaborators means that almost always
the semantic editor services will be disabled because the program is going to be broken in multiple places
transiently)

this is tricky because:

- some edits might be conflicting -- solve this with "conflict holes"

- adding cut/paste or delete/restore allows for degenerate programs (cycles, multiple parents, etc.)

- since we are commutative, we solve both synchronus and async collaborative editing

- and this resolves issues around merges and conflicts

- contribution of this paper is to solve these problems from type-theoretic first principles:

- ...

- Hazel

\subsection{Contributions and Paper Organization}%
\label{sec:Contributions and Paper Organization}