### 2024-07-11 discussion-and-conclusion.tex

\section{Discussion and Conclusion}%
\label{sec:Discussion and Conclusion}

Contribution 1 and Contribution 2 together result in a typed collaborative structure calculus
where, uniquely, all edits, including code relocations that stymie existing approaches, commute and where there are no semantic gaps: all possible editor states, including editor states with various kinds of unresolved conflicts, are semantically meaningful.

In practice, it may be helpful to garbage collect orphaned vertices once there is consensus across collaborators that the deletion is permanent, but we do not consider this consensus protocol formally in this paper. In an open-ended collaboration scenario (where the set of collaborators is not known, e.g. on GitHub), we simply retain all vertices.

TODO: leaves

TODO: our model supports treating these as a cons-list of characters

TODO: (Place somewhere) This move semantics gives us a richer structure than
when treating code as a list of lines.  We exploit
this in Section~REF:TODO in order merge edits that involve moving code.

TODO: Traditional diff: no relation (and indent might change)

TODO: memory usage

TODO: cache eviction algorithm

TODO: Finally, note that through we present several complex scenarios, this is merely for presentation.
In practice, these complex scenarios occur less frequently than portrayed here.

\subsection{Variable names, strings, and numbers}%
\label{sub:Variable names, strings, and numbers}

NOTE: we could implement each digit as a separate characters

GUI for string conflicts: use popups