### 2024-07-11 figures/figure-arity.tex

\newcommand{\figureArityContent}{%
\[
  \arraycolsep=0pt
  \begin{array}{ll}
    \multicolumn{2}{l}{\arityOp : \K \rightarrow \wp(\P \times \{Exp, Pat, Typ\})} \\
    \hline
    \arity{\Root}={} & \left\{ (\Root, Exp) \right\} \\
    \arity{\PatVar(x)}={} & \left\{ \right\} \\
    \arity{\ExpVar(x)}={} & \left\{ \right\} \\
    \arity{\ExpLam}={} & \left\{ (\LamParam, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
    \arity{\ExpApp}={} & \left\{ (\AppFun, Exp), (\AppArg, Exp) \right\} \\
    \arity{\ExpPlus}={} & \left\{ (\PlusLeft, Exp), (\PlusRight, Exp) \right\} \\
    \arity{\ExpTimes}={} & \left\{ (\PlusLeft, Exp), (\PlusRight, Exp) \right\} \\
    \arity{\ExpNum(n)}={} & \left\{ \right\} \\
    \arity{\TypArrow}={} & \left\{ (\ArrowArg, Typ), (\ArrowResult, Typ) \right\} \\
    \arity{\TypNum}={} & \left\{ \right\} \\
  \end{array}
\]%
}

\newcommand{\figureArity}{%
\begin{figure}
\figureArityContent
\caption{Constructors, Indexes and Arity}
\Description{This figure shows the mapping from constructors to their corresponding indexes and arity. For each constructor, such as \ExpLam or \ExpApp, the arity is defined in terms of the parts that make up the constructor. For example, \ExpLam has three parts: a pattern (\LamParam), a type (\mathtt{Type}), and a body (\mathtt{Body}), each with their corresponding expression, pattern, or type.}
\label{fig:Constructors, Indexes and Arity}
\end{figure}%
}