### 2024-07-11 grove-paper.tex

% !TEX program = lualatex

\documentclass[acmsmall,dvipsnames,10pt,review,anonymous,nonacm]{acmart}\settopmatter{printfolios=true}

\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{POPL}
\acmArticle{1}
\acmYear{2025}

\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
\setcopyright{none}

\bibliographystyle{ACM-Reference-Format}
%\citestyle{acmauthoryear}
\citestyle{acmnumeric}

%%% Spacing hacks
\usepackage{titlesec}
\titlespacing*{\section} {0pt}{6pt}{2pt}
% \titlespacing*{\subsection} {0pt}{6pt}{2pt}
\titlespacing*{\subsubsection}{0pt}{6pt}{10pt}
\titlespacing*{\paragraph} {0pt}{4pt}{10pt}
% \titlespacing*{\subparagraph} {12pt}{5pt}{10pt}

\usepackage{enumitem}
\usepackage{listings}%
\lstloadlanguages{ML}
\lstset{tabsize=2, 
basicstyle=\small\ttfamily, 
% keywordstyle=\sffamily,
commentstyle=\itshape\ttfamily\color{gray}, 
stringstyle=\ttfamily\color{purple},
mathescape=false,escapechar=\#,
numbers=none, numberstyle=\scriptsize\color{gray}\ttfamily, language=ML, showspaces=false,showstringspaces=false,xleftmargin=15pt, 
morekeywords={string, float, int},
classoffset=0,belowskip=\smallskipamount, aboveskip=\smallskipamount
}
\let\li\lstinline
\usepackage{todonotes}

% TODO: consolidate all the macros into here
\input{config/macros}

\input{config/config-tex}
\input{config/config-tikz}
\input{config/config-glyphs}
\input{config/config-definitions}
\input{config/config-graphs}

% figures
\input{figures/figure-simple}
\input{figures/figure-wrap-move}
\input{figures/figure-different-parts-nested-parts}
\input{figures/figure-multi-child}
\input{figures/figure-multi-parent}
\input{figures/figure-cycle}
\input{figures/figure-disconnect}
\input{figures/figure-decomposition}
\input{figures/figure-term-syntax}
\input{figures/figure-arity}
\input{figures/figure-decomposition-def}
\input{figures/figure-user-actions}

% marking
\input{symbols}

\begin{document}

%% Title information
\title[Grove]{Grove: A Bidirectionally Typed Collaborative Structure Editor Calculus}
%\subtitle{Subtitle}

%% Author information
\author{Michael D. Adams}
\orcid{0000-0003-3160-6972}

\author{Eric Griffis}
\orcid{0000-0003-1693-6172}

\author{Sundara Vishnu Satish}
\orcid{TODO}

\author{Thomas J. Porter}
\orcid{TODO}

\author{Cyrus Omar}
\orcid{0000-0003-4502-7971}
\affiliation{
  %\position{Assistant Research Scientist}
  \department[0]{Computer Science and Engineering}
  \department[1]{Electrical Engineering and Computer Science}
  \department[2]{College of Engineering}
  \institution{University of Michigan}
  \streetaddress{Bob and Betty Beyster Building, 2260 Hayward Street}
  \city{Ann Arbor}
  \state{MI}
  \postcode{48109-2121}
  \country{USA}
}

% Abstract
\input{abstract}

% 2012 ACM Computing Classification System (CSS) concepts
% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
% \begin{CCSXML}
% <ccs2012>
% <concept>
% <concept_id>10011007.10011006.10011008</concept_id>
% <concept_desc>Software and its engineering~General programming languages</concept_desc>
% <concept_significance>500</concept_significance>
% </concept>
% <concept>
% <concept_id>10003456.10003457.10003521.10003525</concept_id>
% <concept_desc>Social and professional topics~History of programming languages</concept_desc>
% <concept_significance>300</concept_significance>
% </concept>
% </ccs2012>
% \end{CCSXML}
% \ccsdesc[500]{Software and its engineering~General programming languages}
% \ccsdesc[300]{Social and professional topics~History of programming languages}
% End of generated code

%% Keywords
% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission

\maketitle

\input{introduction}
\input{grove-by-example}
\input{formalism}
\input{marking-system}
\input{implementation}
\input{related-work}
\input{discussion-and-conclusion}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
  %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Appendix
\appendix

%% Appendixes that will be published  go **BEFORE** the bibliography and count towards the page count

%% Bibliography
\bibliography{grove-paper}

%% Temporary appendixes that will not be published go **AFTER** the bibliography and do not count towards the page count

\clearpage % Put these appendixes on separate pages so we can easily remove them from the document.  Also ensure all figures have been placed.

% \input{appendix-A-graph-sequences}
\input{appendix/appendix-A-formalism}  
% \input{appendix/appendix-B-marking-system}
\input{appendix/appendix-B-typechecking}
% \input{appendix/appendix-C-type-inference}

\end{document}